---
title: "poli 198 - working code"
author: Sally Rong
output: html_notebook
date: 4/19/21
---

```{r}
# rm(list=ls()) 
# setwd("~/Desktop/POLI 198")
library(tidyverse)
library(dplyr)
```


# Read data in
```{r}
cand2000 <- read.csv("CANDIDATOS_PREFEITO_2000.csv")
  sum(cand2000$CPF_CANDIDATO == "#NULO#") #34 null CPF (60 originally) , all 12 elected/2nd term candidates are color coded matched, some non-elected also matched
cand2004 <- read.csv("CANDIDATOS_PREFEITO_2004.csv")
  sum(cand2004$CPF_CANDIDATO == "#NULO#") # 0 null CPF (7 originally)
cand2008 <- read.csv("CANDIDATOS_PREFEITO_2008.csv")
cand2012 <- read.csv("CANDIDATOS_PREFEITO_2012.csv")
cand2016 <- read.csv("CANDIDATOS_PREFEITO_2016.csv")

votes2000 <- read.csv("TSE_PREFEITO_BR_CANDIDATO_2000.csv")
  sum(votes2000$CPF_CANDIDATO == "#NULO#") #29 null CPF (53 originally) , all 12 elected/2nd term candidates are color coded matched, some non-elected also matched
votes2004 <- read.csv("TSE_PREFEITO_BR_CANDIDATO_2004.csv")
  sum(votes2004$CPF_CANDIDATO == "#NULO#") # 0 null CPF (7 originally), MATCHED!!
votes2008 <- read.csv("TSE_PREFEITO_BR_CANDIDATO_2008.csv")
votes2012 <- read.csv("TSE_PREFEITO_BR_CANDIDATO_2012.csv")
votes2016 <- read.csv("TSE_PREFEITO_BR_CANDIDATO_2016.csv")
#Votes data doesn't have SIGLA_UE (aka municipality) variable, need to verify that the votes data is at municipal level.

## check if there's duplicates, defined by CPF, num_turno
# dup2000 <- votes2000[duplicated(votes2000$CPF_CANDIDATO),] #gives me # duplicate rows
# dim(votes2000[duplicated(votes2000$CPF_CANDIDATO),])[1] # gives me number # duplicates 
# 
# dup2004 <- votes2004[duplicated(votes2004$CPF_CANDIDATO),]
# dim(votes2004[duplicated(votes2004$CPF_CANDIDATO),])[1]
```

# look at Missing CPF's
```{r}
cand2000missing <- cand2000[(cand2000$CPF_CANDIDATO == "#NULO#"),]
votes2000missing <- votes2000[(votes2000$CPF_CANDIDATO == "#NULO#"),]
```

```{r}
length(unique(cand2004$CPF_CANDIDATO))
length(unique(votes2004$CPF_CANDIDATO)) # Notice, there are 15467 unique vs 15561 observations in votes2004 dataset.There discrepancy probably are the runoff elections.
# Votes data appears to be recorded at municipal level; STEP 8 not necessary

length(unique(cand2000$CPF_CANDIDATO))
length(unique(votes2000$CPF_CANDIDATO))
length(unique(cand2008$CPF_CANDIDATO))
length(unique(votes2008$CPF_CANDIDATO))
length(unique(cand2012$CPF_CANDIDATO))
length(unique(votes2012$CPF_CANDIDATO))
length(unique(cand2016$CPF_CANDIDATO))
length(unique(votes2016$CPF_CANDIDATO))

```


# MERGE CANDIDATE AND VOTES DATASETS
```{r}
df2000 <- merge(cand2000, votes2000, by = c("ANO_ELEICAO", "CPF_CANDIDATO"))
sum(df2000$CPF_CANDIDATO == "#NULO#") # 986 null CPF's in df2000 (3180 originally)
df2004 <- merge(cand2004, votes2004, by = c("ANO_ELEICAO", "CPF_CANDIDATO"))
sum(df2004$CPF_CANDIDATO == "#NULO#") # 0 null CPF's in df2004 (49 originally)
df2008 <- merge(cand2008, votes2008, by = c("ANO_ELEICAO", "CPF_CANDIDATO"))
df2012 <- merge(cand2012, votes2012, by = c("ANO_ELEICAO", "CPF_CANDIDATO"))
df2016 <- merge(cand2016, votes2016, by = c("ANO_ELEICAO", "CPF_CANDIDATO"))

## WE NEED TO DECIDE LATER IF WE SHOULD KEEP THE .X AND .Y COLUMNS
```


```{r}
# STEP 8
df2000 <- df2000 %>%
  group_by(SIGLA_UE) %>%    
  mutate(total_mun_votes = sum(QTDE_VOTOS))
length(unique(df2000$total_mun_votes)) # sanity check
# [1] 4263
length(unique(df2000$SIGLA_UE))
# [1] 5015
df2004 <- df2004 %>%
  group_by(SIGLA_UE) %>%    
  mutate(total_mun_votes = sum(QTDE_VOTOS))
length(unique(df2004$total_mun_votes))
# [1] 4574
df2008 <- df2008 %>%
  group_by(SIGLA_UE) %>%    
  mutate(total_mun_votes = sum(QTDE_VOTOS))
df2012 <- df2012 %>%
  group_by(SIGLA_UE) %>%    
  mutate(total_mun_votes = sum(QTDE_VOTOS))
df2016 <- df2016 %>%
  group_by(SIGLA_UE) %>%    
  mutate(total_mun_votes = sum(QTDE_VOTOS))

# STEP 9 - PROP_VOTES variable: Proportion of votes each candidate received
df2000$PROP_VOTES <- df2000$QTDE_VOTOS/df2000$total_mun_votes
df2004$PROP_VOTES <- df2004$QTDE_VOTOS/df2004$total_mun_votes
df2008$PROP_VOTES <- df2008$QTDE_VOTOS/df2008$total_mun_votes
df2012$PROP_VOTES <- df2012$QTDE_VOTOS/df2012$total_mun_votes
df2016$PROP_VOTES <- df2016$QTDE_VOTOS/df2016$total_mun_votes


sigla85219 <- subset(df2000, SIGLA_UE == 85219)
```

# STEP 5 -  INCUMBENT VARIABLE
```{r}
unique(df2004$DESC_SIT_TOT_TURNO.x)
  ## what does #NULO# mean in the DESC_SIT_TOT_TURNO variable?

# Create subsets where DESC_SIT_TOT_TURNO = "ELEITO"
#df2004_winners will be a subset of df2004 where DESC_SIT_TOT_TURNO = "ELEITO"
df2000_winners <- subset(df2000, DESC_SIT_TOT_TURNO.x == "ELEITO")
df2004_winners <- subset(df2004, DESC_SIT_TOT_TURNO.x == "ELEITO")
df2008_winners <- subset(df2008, DESC_SIT_TOT_TURNO.x == "ELEITO")
df2012_winners <- subset(df2012, DESC_SIT_TOT_TURNO.x == "ELEITO")
df2016_winners <- subset(df2016, DESC_SIT_TOT_TURNO.x == "ELEITO")

df2004$INCUMBENT <- df2004$CPF_CANDIDATO %in% df2000_winners$CPF_CANDIDATO # generates T/F values
df2004$INCUMBENT <- as.integer(df2004$INCUMBENT == "TRUE")  #turns T/F values into 1,0 (dummy var)

df2008$INCUMBENT <- df2008$CPF_CANDIDATO %in% df2004_winners$CPF_CANDIDATO 
df2008$INCUMBENT <- as.integer(df2008$INCUMBENT == "TRUE")

df2012$INCUMBENT <- df2012$CPF_CANDIDATO %in% df2008_winners$CPF_CANDIDATO
df2012$INCUMBENT <- as.integer(df2012$INCUMBENT == "TRUE")

df2016$INCUMBENT <- df2016$CPF_CANDIDATO %in% df2012_winners$CPF_CANDIDATO
df2016$INCUMBENT <- as.integer(df2016$INCUMBENT == "TRUE")
```

# STEP 6 - INCUMBENT_WINS Variable
```{r}
df2004$INCUMBENT_WINS <- c(df2004$INCUMBENT == 1 & df2004$DESC_SIT_TOT_TURNO.x == "ELEITO" )
df2004$INCUMBENT_WINS <- as.integer(df2004$INCUMBENT_WINS == "TRUE")


df2008$INCUMBENT_WINS <- c(df2008$INCUMBENT == 1 & df2008$DESC_SIT_TOT_TURNO.x == "ELEITO" )
df2008$INCUMBENT_WINS <- as.integer(df2008$INCUMBENT_WINS == "TRUE")

df2012$INCUMBENT_WINS <- c(df2012$INCUMBENT == 1 & df2012$DESC_SIT_TOT_TURNO.x == "ELEITO" )
df2012$INCUMBENT_WINS <- as.integer(df2012$INCUMBENT_WINS == "TRUE")

df2016$INCUMBENT_WINS <- c(df2016$INCUMBENT == 1 & df2016$DESC_SIT_TOT_TURNO.x == "ELEITO" )
df2016$INCUMBENT_WINS <- as.integer(df2016$INCUMBENT_WINS == "TRUE")
```

# Create TERM-LIMITED Variable
```{r}
# If for any candidate, in previous election, the Incumbent_wins variable == 1 , then municipality is term-limited in this election


# term_lim2004 <- subset(df2000, INCUMBENT_WINS == 1)
# df2004$TERM_LIMITED <- df2004$CPF_CANDIDATO %in% term_lim2004$CPF_CANDIDATO
# df2004$TERM_LIMITED <- as.integer(df2004$TERM_LIMITED == "TRUE")
# sum(df2004$TERM_LIMITED == 1)

term_lim2008 <- subset(df2004, INCUMBENT_WINS == 1)
df2008$TERM_LIMITED <- df2008$CPF_CANDIDATO %in% term_lim2008$CPF_CANDIDATO
df2008$TERM_LIMITED <- as.integer(df2008$TERM_LIMITED == "TRUE")
sum(df2008$TERM_LIMITED == 1)

term_lim2012 <- subset(df2008, INCUMBENT_WINS == 1)
df2012$TERM_LIMITED <- df2012$CPF_CANDIDATO %in% term_lim2012$CPF_CANDIDATO
df2012$TERM_LIMITED <- as.integer(df2012$TERM_LIMITED == "TRUE")
sum(df2012$TERM_LIMITED == 1)

term_lim2016 <- subset(df2012, INCUMBENT_WINS == 1)
df2016$TERM_LIMITED <- df2016$CPF_CANDIDATO %in% term_lim2016$CPF_CANDIDATO
df2012$TERM_LIMITED <- as.integer(df2012$TERM_LIMITED == "TRUE")
sum(df2016$TERM_LIMITED == 1)
```

# Create a NUM_CAND variable to display how many candidates are running in a municipality
```{r}
df2000 <- df2000 %>%
  group_by(SIGLA_UE) %>%    
  mutate(NUM_CAND = length(unique(CPF_CANDIDATO)))
unique(df2000$NUM_CAND)
### THERE ARE DUPLICATES!!
# df2000 %>% filter(NUM_CAND == 15)

df2004 <- df2004 %>%
  group_by(SIGLA_UE) %>%    
  mutate(NUM_CAND = length(unique(CPF_CANDIDATO)))
df2008 <- df2008 %>%
  group_by(SIGLA_UE) %>%    
  mutate(NUM_CAND = length(unique(CPF_CANDIDATO)))
df2012 <- df2012 %>%
  group_by(SIGLA_UE) %>%    
  mutate(NUM_CAND = length(unique(CPF_CANDIDATO)))
df2016 <- df2016 %>%
  group_by(SIGLA_UE) %>%    
  mutate(NUM_CAND = length(unique(CPF_CANDIDATO)))
```

# Create PROP_VOTES_WINNER variable
```{r}
df2000 <- df2000 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_WINNER = max(PROP_VOTES))
df2004 <- df2004 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_WINNER = max(PROP_VOTES))
df2008 <- df2008 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_WINNER = max(PROP_VOTES))
# DOUBLE CHECKING 
#df2008 %>%
#  filter(SIGLA_UE == 2046) %>%    
#  mutate(PROP_VOTES_WINNER = max(PROPORTION_VOTES))
df2012 <- df2012 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_WINNER = max(PROP_VOTES))
df2016 <- df2016 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_WINNER = max(PROP_VOTES))
```

# Create PROP_VOTES_RUNNERUP variable
```{r}
df2000 <- df2000 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_RUNNERUP = sort(PROP_VOTES, TRUE)[2])
df2004 <- df2004 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_RUNNERUP = sort(PROP_VOTES, TRUE)[2])
df2008 <- df2008 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_RUNNERUP = sort(PROP_VOTES, TRUE)[2])
# DOUBLE CHECKING 
# df2008 %>% filter(SIGLA_UE == 2046)
df2012 <- df2012 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_RUNNERUP = sort(PROP_VOTES, TRUE)[2])
df2016 <- df2016 %>%
  group_by(SIGLA_UE) %>%    
  mutate(PROP_VOTES_RUNNERUP = sort(PROP_VOTES, TRUE)[2])
```

# Create WINNER_PARTY variable, which lists the party of the winner for that election (grouping by municipality)
```{r}
# I first created another replacing dataframe of all the parties of winners for each municipality
winner_replace_df2000 <- df2000 %>%
  group_by(SIGLA_UE) %>%    
  filter(DESC_SIT_TOT_TURNO.x == "ELEITO") %>%
  select(SIGLA_UE, WINNER_PARTY = SIGLA_PARTIDO.x)
# Then, I join that replacing dataframe w/ the original
df2000 <- df2000 %>%
  group_by(SIGLA_UE) %>% 
  left_join(replace_values, by = "SIGLA_UE")

# 2004
winner_replace_df2004 <- df2004 %>%
  group_by(SIGLA_UE) %>%    
  filter(DESC_SIT_TOT_TURNO.x == "ELEITO") %>%
  select(SIGLA_UE, WINNER_PARTY = SIGLA_PARTIDO.x)
df2004 <- df2004 %>%
  group_by(SIGLA_UE) %>% 
  left_join(replace_values, by = "SIGLA_UE")

# 2008
winner_replace_df2008 <- df2008 %>%
  group_by(SIGLA_UE) %>%    
  filter(DESC_SIT_TOT_TURNO.x == "ELEITO") %>%
  select(SIGLA_UE, WINNER_PARTY = SIGLA_PARTIDO.x)
df2008 <- df2008 %>%
  group_by(SIGLA_UE) %>% 
  left_join(replace_values, by = "SIGLA_UE")

# 2012
winner_replace_df2012 <- df2012 %>%
  group_by(SIGLA_UE) %>%    
  filter(DESC_SIT_TOT_TURNO.x == "ELEITO") %>%
  select(SIGLA_UE, WINNER_PARTY = SIGLA_PARTIDO.x)
df2012 <- df2012 %>%
  group_by(SIGLA_UE) %>% 
  left_join(replace_values, by = "SIGLA_UE")

# 2016
winner_replace_df2016 <- df2016 %>%
  group_by(SIGLA_UE) %>%    
  filter(DESC_SIT_TOT_TURNO.x == "ELEITO") %>%
  select(SIGLA_UE, WINNER_PARTY = SIGLA_PARTIDO.x)
df2016 <- df2016 %>%
  group_by(SIGLA_UE) %>% 
  left_join(replace_values, by = "SIGLA_UE")
```

# Create 'runoff' variable, a dummy variable of municipalities that went to a 2nd election.
```{r}
df2000$runoff <- as.integer(df2000$DESC_SIT_TOT_TURNO.x == "2º TURNO")
df2004$runoff <- as.integer(df2004$DESC_SIT_TOT_TURNO.x == "2º TURNO")
df2008$runoff <- as.integer(df2008$DESC_SIT_TOT_TURNO.x == "2º TURNO")
df2012$runoff <- as.integer(df2012$DESC_SIT_TOT_TURNO.x == "2º TURNO")
df2016$runoff <- as.integer(df2016$DESC_SIT_TOT_TURNO.x == "2º TURNO")
```


# Merging datasets
```{r}
df2000 <- df2000 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))
df2004 <- df2004 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))
df2008 <- df2008 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))
df2012 <- df2012 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))
df2016 <- df2016 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))

df2000 <- df2000 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))
df2004 <- df2004 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))
df2008 <- df2008 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))
df2012 <- df2012 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))
df2016 <- df2016 %>% mutate(CPF_CANDIDATO = as.character(CPF_CANDIDATO))

df2000 <- df2000 %>% mutate(NUM_TITULO_ELEITORAL_CANDIDATO = as.character(NUM_TITULO_ELEITORAL_CANDIDATO))
df2004 <- df2004 %>% mutate(NUM_TITULO_ELEITORAL_CANDIDATO = as.character(NUM_TITULO_ELEITORAL_CANDIDATO))
df2008 <- df2008 %>% mutate(NUM_TITULO_ELEITORAL_CANDIDATO = as.character(NUM_TITULO_ELEITORAL_CANDIDATO))
df2012 <- df2012 %>% mutate(NUM_TITULO_ELEITORAL_CANDIDATO = as.character(NUM_TITULO_ELEITORAL_CANDIDATO))
df2016 <- df2016 %>% mutate(NUM_TITULO_ELEITORAL_CANDIDATO = as.character(NUM_TITULO_ELEITORAL_CANDIDATO))

mergedData <- rbind(df2000, df2004, df2008, df2012, df2016)
write.csv(mergedData, "mergedData.csv")
```

Questions
What to do with duplicates? (run-offs elections)


